
\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

\usepackage[backend=bibtex]{biblatex}
\addbibresource{retinopathy.bib}

\begin{document}

\section*{Personal Information}
Name: \hfill Jordi de la Torre Gallart\\
ID card number: \hfill 39.724.316-G\\
Phone: \hfill +34 676 533 655\\
e-mail: \hfill jordi.delatorre@gmail.com\\
web: \hfill jorditg.github.io\\

\section*{Studies} 

{\sl M.Sc. Chemical Engineering}. URV. Tarragona.Spain \hfill (1998)\\
{\sl M.D. Business Administration}. UOC. Barcelona.  Spain\hfill (2011)\\
{\sl B.Sc. Computer Science \hfill (2004)\\ }	
{\sl Ph.D. Computer Science \hfill (2019)\\ }	

\section*{Research areas and publications}
	
	\begin{enumerate}
		\item Design of automatic classifiers based on deep neural networks able to reach ophthalmologist performance level.
		
		\fullcite{jdelatorre2016}
		
		\item Study of the usage of Quadratic Weighted Kappa index as a Deep Learning Loss Function for the optimization of ordinal regression problems.
		
		\fullcite{delatorre2017} Impact Factor: 1.952 (Q2)
		
		\item Design of a generalized model for the interpretation of results reported by deep learning classifiers.
		
		\fullcite{de2017deep} Accepted for publication in Neurocomputing. Impact Factor: 3.241 (Q1)
		
		\item Design of a method for compressing feature space internal representations of deep learning models.
		
		\fullcite{delatorre_ica_2018}
		
	\end{enumerate}
	
\section*{Conferences}

	\begin{enumerate}
\item 19st International Conference of the Catalan Association for Artificial Intelligence, Barcelona, October 2016 (Poster presentation)
\item 4th URV Doctoral Workshop in Computer Science and Mathematics, Tarragona, 2017 (Oral presentation)
	\end{enumerate}
	%\printbibliography
	
\end{document}



