\documentclass{elsarticle}

\begin{document}
	
\section*{Authors' Biographies}

\subsection*{Jordi de la Torre}
\textbf{Jordi de la Torre} is a Machine Learning and Artificial Intelligence Ph.D. student in Universitat Rovira i Virgili, Tarragona, Spain. His supervisors are Dr. Aida Valls and Dr. Domenec Puig. He has a multidisciplinary background. He received a M.S. in Chemical Engineering in 1998 and a B.S. in Computer Science in 2007 by Universitat Rovira i Virgili, and a B.S. and M.S. in Business Administration in 2011 by the Universitat Oberta de Catalunya, Barcelona. His main research area is Deep Learning applied to Biology. 

\subsection*{Dr. Aida Valls}
\textbf{Dr. Aida Valls} is a lecturer at the Department of Computer Science and Mathematics in Universitat Rovira i Virgili (URV). She received the PhD in Artificial Intelligence from the Polytechnical University of Catalonia in 2002. Her research interests include multiple criteria decision making, recommender systems, data mining and machine learning. Her work is mainly focused on the treatment of linguistic and semantic information. She has participated in several Spanish and EU research projects, with applications in Tourism, Environment Risk Management and Health Care. She is the author of 50 papers in international journals and  more than 70 papers in conferences. In Google Shcolar Dr. Valls has a h-index of 24 with 2270 citations. Dr. Valls has been the supervisor of 4 PhD Thesis and is now supervising 4 PhD students. She is currently the Head of the PhD Program in Computer Science and Mathematics of Security at URV and the Coordinator of Mobility of Computer Science studies at URV.

\subsection*{Dr. Domenec Puig}
\textbf{Dr. Domenec Puig} received a degree in computer science and a doctorate from the Polytechnic University of Catalonia, Barcelona, ​​Spain, in 1992 and 2004, respectively. In 1992, he joined the Department of Computer Engineering and Mathematics of the University Rovira i Virgili (URV), Tarragona, Spain, where he is Full Professor since June 2017. He currently serves as the director of the School of Engineering of the URV since 2013. Since 2006, he is the principal investigator of the Intelligent Robotics and Computer Vision research group of the URV. His research interests include: image processing, pattern recognition, computer vision, medical image analysis, machine learning, perception models, scene analysis, mobile robotics and socially assistive robotics. Dr. Domenec Puig has participated as a researcher in more than 35 research projects related to his topics of interest, funded by public administrations (autonomous, state and European), among which he has been the principal investigator in 22 projects. In addition, he has participated in more than 25 contracts, agreements or R\&D+i projects with public or private administrations or entities, being the principal investigator in 23. Dr. Domenec Puig has participated in more than 150 publications among specialized journals, book chapters and conferences. Among these publications, more than 50 in journals indexed in the Journal Citation Reports of the ISI Web of Knowledge. Dr. Domenec Puig has supervised 8 doctoral theses presented in recent years, related to their topics of interest in Robotics, Computer Vision, Image Processing and Pattern Recognition. Dr. Domenec Puig has participated in numerous international conferences, acting as a speaker, reviewer of works and chair of sessions. He has also been a reviewer of national and international competitive research projects, as well as reviewer of articles in prestigious international journals.

\subsection*{Dr. Pedro Romero-Aroca}
\textbf{Dr. Pedro Romero-Aroca} is Director of the Ophthalmic Department of Hospital Universitari de Sant Joan de Reus, Catalonia, Spain, Associate Professor of Ophthalmology of Universitat Rovira i Virgili. Research deals on diabetic retinopathy (DR), DR epidemiology, telemedicine, diabetic macular edema etiology and its oxidative and inflammatory basis.

\end{document}